Individual heterogeneity in the functional network topography of medial prefrontal

### 2019-09-12 Manuscript.Rmd
Sean revisions up to results, completed

All network analyses were performed using the igraph package [v. 1.1.2; https://igraph.org/r/; @G.2006] in R [v. 3.4.1; https://www.r-project.org/; @RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. All correlation values were transformed using Fisher's r to z. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Edges mostly consisted of positive correlations (mean proportion positive = `r mPos`, sd = `r sdPos`). We chose not to threshold this network, as SP is well equipped to operate on complete (i.e. fully-connected) graphs [@chung1997]. However, our results are unchanged even if we cut all non-significant correlations from the weight matrices. Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Ensuring that all edges were positive facilitated the construction of the graph Laplacian (see below), which requires all off-diagonal elements to have the same sign by design. We generated and analyzed network weight matrices at four levels: (1) for each subject's full concatenated dataset (up to 4800 TRs); (2) on each step of a sliding window analysis (see Partition Evaluation for more details); (3) for the concatenated time series for the two runs on each day (up to 2400 TRs); and (4) for each run separately (up to 1200 TRs).

@@

In order to evaluate the validity of the resulting partitionings across community-detection methods, we also estimated network communities using the more traditional approach of modularity maximization [@Garcia2017], based on the algorithm from Clauset et al. [-@Clauset2004]. The method heuristically iterates through many possible combinations of vertices, and selects the partitioning that maximizes the within-community edge weights, relative to a random network containing the same number of edges and communities. Unlike SP, modularity can fractionate a network into more than two communities. Agreement between the partitions provided by the bounded and unbounded community detection methods would suggest the results are not distorted by the restriction of SP to binary partitionings.

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject. SP and Modularity have been previously found to show underfitting and overfitting tendencies, respectively, in their community detection performance in a diverse set of network types [@Ghasemian2018], so alignment between the two algorithms would increase our confidence in the validity of the resulting partitionings. Next, we compared the subject-level SP partitionings across individuals, and calculated the mean pairwise ARI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there were differences in overall agreement within these regions by performing a paired permutation analysis. We randomized the sign of the difference in pairwise ARI across individuals between mPFC and PCC 5000 times, and computed the mean of these differences across all comparisons on each iteration. The empirical difference in mean pairwise ARI between regions was then evaluated against this permuted distribution.

To estimate the degree of variability of partitionings per individual we performed a sliding window analysis (20 min windows, 1 min increments, median number of windows per subject = 37, range = 35 - 37), comparing each window's resulting partitioning against the partitioning derived from the subject's whole data set. A 20-min window has previously been found to approximate an individual's asymptote for the stability of community features in brain networks [@Gordon2017]. We assessed whether the magnitude of the Fiedler vector value for a given vertex (for the full subject-level data set) was associated with the stability of that vertex’s sub-network assignment across time windows. To do this, we fit a mixed effects logistic regression model, in which the dependent variable was the proportion of times each vertex participated in the DN community across windows, and the explanatory variables included a random effect of subject and a fixed effect of the Fiedler vector value for that vertex (derived from their full time series). Based on this significant relationship, we identified a threshold Fiedler vector value for each subject, such that empirical above-threshold vertices were persistently associated with either DN or non-DN more than 99% of the time.

@@

Next we asked whether the location of the DN and non-DN communities was guided by sulcal morphology. Based on previous reports of individual alignment of DN within sulci in ventral mPFC [@Lopez2019], we sub-divided our search space into three regions: vmPFC, which matched their ROI (areas 25, s32, a24, 10v, 10r, p32, and OFC); superior mPFC (sup-mPFC), encompassing the remaining dorsal areas in our mPFC space; and all of the PCC search space. We used each subject's curvature maps from the HCP (transformed to fs_LR 32k space), in which cortical depth is quantified by negative numbers for sulci and positive numbers for gyri. For each individual, we computed the correlation (Spearman) between curvature and the unthresholded Fiedler vector values in each region and hemisphere separately (6 correlations per individual). Since the DN is indicated by positive FV values, positive correlations meant that the DN was more likely to be contained in sulci, with non-DN housed in gyri. Finally, we collected all individual correlations for each combination of hemisphere and region, and performed a one-sample t-test on each set to determine whether correlations were significantly different than 0 in our group (6 tests total).

@@

We next sought to estimate the variability of individual partitionings over time. We did so by performing a sliding window analysis on each subject's full time series (20 min windows shifting by 1 min). We compared the partitioning derived from each window with the partitioning computed using the entire time series (Figure 4). The significant level of sequential overlap was meant to promote highly similar partitionings among nearby windows, so that we could identify nodes that were highly variable where discrepancies occurred. This also allowed us to determine whether these variable nodes followed a specific spatial structure.

@@

We next tested whether the continuous-valued Fiedler vector (before binarization into discrete communities) carried information about the stability of individual nodes. There is precedent in the literature for the idea that the magnitude (and not just the sign) of the Fiedler vector values conveys important information about the role of each node in the network [@Gkantsidis2003; @Tian2018]. Therefore, we tested whether the magnitude of the eigenvector values was associated with the stability of nodes over time. Specifically, we estimated the proportion of DN affiliations per node as a function of Fiedler vector values, using a logistic mixed effects model (Figure 4). We stress that the stability expected from our highly overlapping windows added relevance to the observed variability. The model identified a positive significant relationship between these features ($\beta$ = `r round(SC_FVvPDN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDN_modelSummary$coefficients[2, 2], digits = 2)`, *p* < 0.001), signifying that vertices with higher absolute Fiedler vector values were more persistent in their relationship with their corresponding community over time.

@@

The relatively high inter-individual variability seen in the aggregate map could reflect at least three factors: (1) measurement noise, (2) dynamic variation in mPFC network organization, and (3) stable patterns of functional organization that differ across individuals. To arbritrate among these possibilities, we examined the test/re-test reliability of mPFC/PCC community structure across separate days of testing. Insofar as the observed variability reflects individual-specific brain organization, across-day ARI values should be consistently higher within-individual than between individuals (an example comparison for two individuals is provided in Supplemental Figure 2). Figure 6 shows pairwise comparisons among ten example subjects for PCC and mPFC separately (left).

>>>>

All network analyses were performed using the igraph package [v. 1.1.2; https://igraph.org/r/; @G.2006] in R [v. 3.4.1; https://www.r-project.org/; @RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. All correlation values were transformed using Fisher's r to z. This produced a weighted network for each subject, in which the nodes were surface vertices and the edge weights were the correlations among them. Edges mostly consisted of positive correlations (mean proportion positive = `r mPos`, sd = `r sdPos`). We chose not to threshold this network, as SP is well equipped to operate on complete (i.e. fully-connected) graphs [@chung1997]. However, our results are unchanged even if we cut all non-significant correlations from the weight matrices (uncorrected). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Ensuring that all edges were positive facilitated the construction of the graph Laplacian (see below), which requires all off-diagonal elements to have the same sign by design. We generated and analyzed network weight matrices at four levels: (1) for each subject's full concatenated dataset (up to 4800 TRs); (2) on each step of a sliding window analysis (see Partition Evaluation for more details); (3) for the concatenated time series for the two runs on each day (up to 2400 TRs); and (4) for each run separately (up to 1200 TRs).

@@

In order to evaluate the validity of the resulting partitionings across community-detection methods, we also estimated network communities using the more traditional approach of modularity maximization [@Garcia2017], based on the algorithm from Clauset et al. [-@Clauset2004]. The method heuristically iterates through many possible combinations of vertices, and selects the partitioning that maximizes the within-community edge weights, relative to a random network containing the same number of edges and communities. Unlike SP, modularity can fractionate a network into more than two communities. Agreement between the partitions provided by the bounded (SP) and unbounded (modularity) community detection methods would suggest the results are not distorted by the restriction of SP to binary partitionings.

@@

We performed a number of comparisons among partitions. First, we computed the degree of agreement between SP and modularity per subject. SP and Modularity have been previously found to show a tendency for underfitting and overfitting, respectively, in their community detection performance in a diverse set of network types [@Ghasemian2018], so alignment between the two algorithms would increase our confidence in the validity of the resulting partitionings. Next, we compared the subject-level SP partitionings across individuals, and calculated the mean pairwise ARI for the group. We then performed the same evaluation for PCC and mPFC separately, and examined whether there were differences in overall agreement within these regions by performing a paired permutation analysis. We randomized the sign of the difference in pairwise ARI across individuals between mPFC and PCC 5000 times, and computed the mean of these differences across all comparisons on each iteration. The empirical difference in mean pairwise ARI between regions was then evaluated against this permuted distribution.

To estimate the degree of variability of partitionings per individual we performed a sliding window analysis (20 min windows, 1 min increments, median number of windows per subject = 37, range = 35 - 37), comparing each window's resulting partitioning against the partitioning derived from the subject's whole data set. A 20-min window has previously been found to approximate asymptotic stability of community features in brain networks as time series are lengthened [@Gordon2017]. We assessed whether the magnitude of the Fiedler vector value for a given vertex (for the full subject-level data set) was associated with the stability of that vertex’s sub-network assignment across time windows. To do this, we fit a mixed effects logistic regression model, in which the dependent variable was the proportion of times each vertex participated in the DN community across windows, and the explanatory variables included a random effect of subject and a fixed effect of the Fiedler vector value for that vertex (derived from their full time series). Based on this significant relationship, we identified a threshold Fiedler vector value for each subject, such that empirical above-threshold vertices were persistently associated with either DN or non-DN more than 99% of the time.

@@

Next we asked whether the location of the DN and non-DN communities was guided by sulcal morphology. Based on previous reports of individual alignment of DN within sulci in ventral mPFC [@Lopez2019], we sub-divided our search space into three regions: vmPFC, which matched their ROI (areas 25, s32, a24, 10v, 10r, p32, and OFC); superior mPFC (sup-mPFC), encompassing the remaining dorsal areas in our mPFC space; and all of the PCC search space. We used each subject's curvature maps from the HCP (transformed to fs_LR 32k space), in which cortical depth is quantified by negative numbers for sulci and positive numbers for gyri. For each individual, we computed the Spearman correlation between curvature and the unthresholded Fiedler vector values in each region and hemisphere separately (6 correlations per individual). Since the DN is indicated by positive FV values, positive correlations meant that the DN was more likely to be contained in sulci, with non-DN housed in gyri. Finally, we collected all individual correlations for each combination of hemisphere and region, and performed a one-sample t-test on each set to determine whether correlations were significantly different than 0 in our group (6 tests total).

@@

We next sought to estimate the variability of individual partitionings over time. We did so by performing a sliding window analysis on each subject's full time series (20 min windows shifting by 1 min). We compared the partitioning derived from each window with the partitioning computed using the entire time series (Figure 4). The significant level of sequential overlap was meant to promote highly similar partitionings among nearby windows, so that we could identify nodes that were highly variable where discrepancies occurred (i.e. highly variable nodes that changed their partition membership with minor perturbations to the windowed data). *This also allowed us to determine whether these variable nodes followed a specific spatial structure.*

@@

We next tested whether the continuous-valued Fiedler vector (before binarization into discrete communities) carried information about the stability of individual nodes. There is precedent in the literature for the idea that the magnitude (and not just the sign) of the Fiedler vector values conveys important information about the role of each node in the network [@Gkantsidis2003; @Tian2018]. Therefore, we tested whether the magnitude of the eigenvector values was associated with the stability of nodes over time. Specifically, we estimated the proportion of DN affiliations per node as a function of Fiedler vector values, using a logistic mixed effects model (Figure 4). The model identified a positive significant relationship between these features ($\beta$ = `r round(SC_FVvPDN_modelSummary$coefficients[2,1], digits = 2)`, SE = `r round(SC_FVvPDN_modelSummary$coefficients[2, 2], digits = 2)`, *p* < 0.001), signifying that vertices with higher absolute Fiedler vector values were more persistent in their relationship with their corresponding community over time.

@@

The relatively high inter-individual variability seen in the aggregate map could reflect at least three factors: (1) measurement noise, (2) dynamic variation in mPFC network organization, and (3) stable patterns of functional organization that differ across individuals. To arbitrate among these possibilities, we examined the test/re-test reliability of mPFC/PCC community structure across separate days of testing. Insofar as the observed variability reflects individual-specific brain organization, across-day ARI values should be consistently higher within-individual than between individuals (an example comparison for two individuals is provided in Supplemental Figure 2). Figure 6 shows pairwise comparisons among ten example subjects for PCC and mPFC separately (left).
